---
title: "Soil Column ARG Dispersal Workflow"
author: "Schuyler D. smith"
date: "`r Sys.Date()`"
---
<br>
<strong><a href="https://github.com/schuyler-smith/soil_column_ARG_disperal" target="_blank">Project Repository with Data</a></strong>
<br>
<br>
<strong><a href="https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0222470" target="_blank">Investigating the dispersal of antibiotic resistance associated genes from manure application to soil and drainage waters in simulated agricultural farmland systems</a></strong>
<br>
<br>
#Required Pacakges
```{r install-packages, eval=FALSE}
devtools::install_github('schuyler-smith/phylosmith')
devtools::install_github('schuyler-smith/schuylR')
```


```{r required-packages}
library(data.table)
library(phyloseq)
library(phylosmith)
library(ggplot2)
library(vegan)
library(schuylR)
library(gridExtra)
```

#Data Setup

```{r load-data}
phylo_object <- readRDS("../data/soil_column_ARGs.RDS"); phylo_object

phylo_object <- phylosmith::merge_treatments(phylo_object, 'Matrix', 'Treatment')
treatments <- 'Matrix_Treatment'
levels(phylo_object@sam_data$Matrix_Treatment) <- c('Effluent', 'Manure', 'Soil', 'Manure-Treated Effluent', NA, 'Manure-Treated Soil')
phylo_object <- phylosmith::set_treatment_levels(phylo_object, treatment = 'Matrix_Treatment', 
                order = c('Manure', 'Soil', 'Effluent', 'Manure-Treated Soil', 'Manure-Treated Effluent'))

metadata <- fread('../data/metadata.tsv')
gsub('SC.*_A', 'A', metadata$`Sample ID`) == sample_names(phylo_object)
sample_names(phylo_object) <- paste(sapply(strsplit(metadata$`Sample ID`, '_'),  `[`, 1), metadata$Day, sep = '_Day_')
phylo_object
```

```{r initiate environment, echo=FALSE}
knitr::opts_chunk$set(output.lines=6, cache = TRUE, fig.width = 16, fig.height = 8)
phylo_object <- classify_ARG_classes(phylo_object, "Accession", combine = 2)
phylo_object <- classify_ARG_mechanisms(phylo_object, "Accession")
phylo_object@tax_table[,'ARG_Class'][phylo_object@tax_table[,'ARG_Class'] == "Antibiotic Without Defined Classification",] <- 'Unclassified'
phylo_object <- set_treatment_levels(phylo_object, 'Matrix', c('Manure', 'Soil', 'Effluent'))
phylo_object <- set_treatment_levels(phylo_object, 'Treatment', c('Control', 'Manure'))
phylo_object <- set_treatment_levels(phylo_object, 'Day', 'numeric')
phylo_object <- set_sample_order(phylo_object, c('Matrix_Treatment', 'Day'))
phylo_object <- taxa_filter(phylo_object, treatments)
```

```{r separate-systems, echo = FALSE}
manure <- taxa_filter(phylo_object, treatments, subset = 'Manure', frequency = 0)
manure_genes <- taxa_names(manure)
soil <- taxa_filter(phylo_object, treatments, subset = c('Soil', 'Effluent'), frequency = 0)
soil_genes <- taxa_names(soil)
```

#Manure Resistome vs Soil Resistome

```{r manure-genes}
manure
```
```{r soil-genes}
soil
```

##Unique Genes for Each Resistome

```{r unique-control-genes}
unique_control_genes <- unique_taxa(phylo_object, treatments, c('Manure', 'Soil', 'Effluent')); sapply(unique_control_genes, length)
```

```{r class_proportions_manure}
data.frame(taxa_proportions(manure, 'ARG_Class'), row.names = 1)
```

```{r class_proportions_soil}
data.frame(taxa_proportions(soil, 'ARG_Class'), row.names = 1)
```

```{r abundance-comparison}
sum(taxa_sums(taxa_filter(phylo_object, treatment = treatments, subset = 'Manure')))/sum(taxa_sums(taxa_filter(phylo_object, treatment = treatments, subset = 'Soil')))
```

#Phylogeny Profiles

```{r phylogeny-bar-plots, fig.fullwidth=TRUE, warning=FALSE}
g <- phylogeny_profile(conglomerate_samples(phylo_object, treatment = treatments, merge_on = 'Day'), 'ARG_Class', treatments, relative_abundance = TRUE) + 
  labs(fill = 'Associated Anti-Microbial\nResistance Drug Class', x = 'Sampling Day') + 
  guides(fill=guide_legend(ncol=1))
g
```

#Ordination

```{r tsne-relative-abundance, echo=FALSE}
phyloseq_obj = relative_abundance(phylo_object)
treatment = c(treatments)
phyloseq_obj <- taxa_filter(phyloseq_obj, treatment, frequency = 0)
treatment_name <- paste(treatment, collapse = '_')
metadata <- as(access(phyloseq_obj, "sam_data"), "data.frame")
color_count <- length(unique(metadata[[treatment_name]]))
graph_colors <- phylosmith:::create_palette(color_count, "default")
tsne <- Rtsne::Rtsne(vegdist(t(access(phyloseq_obj, "otu_table")), 
  method = "bray"), dims = 2, theta = 0, perplexity = 8)
tSNE1 <- tsne$Y[, 1]
tSNE2 <- tsne$Y[, 2]
ord <- data.table(tSNE1, tSNE2, metadata)
ord <- subset(ord, !is.na(treatment_name))
g <- ggplot(data = ord, aes(tSNE1, tSNE2, group = Matrix_Treatment))
points <- ggplot_build(g)$data[[1]]
ellipse_df <- phylosmith:::CI_ellipse(points, 'group', level = 0.95)
g <- g + geom_polygon(data = ellipse_df, aes(x = x, y = y, group = group), color = 'black', 
                 fill = graph_colors[ellipse_df$group], alpha = 0.3, size = 0.6, linetype = 1)
g <- g + geom_point(aes(fill = Matrix_Treatment), shape = c(21,22,24)[ord[['Matrix']]], 
  color = "black", size = 7, alpha = 1) + scale_fill_manual(values = graph_colors) +
  theme_classic() + theme(aspect.ratio = 1, axis.line.x = element_line(colour = "black", 
    size = 1, linetype = "solid"), axis.line.y = element_line(colour = "black", 
    size = 1, linetype = "solid"), axis.text.x = element_text(size = 10), 
    axis.text.y = element_text(size = 10), axis.title.x = element_text(size = 10, 
      face = "bold"), axis.title.y = element_text(size = 10, 
      face = "bold"), legend.title = element_text(size = 10, face = 'bold'), 
    legend.spacing.x = unit(0.005, 'npc'),
    legend.text = element_text(size = 8), legend.background = element_rect(fill = (alpha = 0))) + 
    labs(x = "t-SNE Dimension 1", y = "t-SNE Dimension 2")
g + guides(fill = guide_legend(override.aes=list(shape=c(21,22,24,22,24)), keywidth=0.4,
                 default.unit = "inch")) + labs(fill = 'Sample Type')
```

#Statistical Tests

```{r Mann-Whitney-test}
wilcox.test(as(manure@otu_table, 'matrix'), as(soil@otu_table, 'matrix'))
wilcox.test(as(taxa_filter(phylo_object, treatments, subset = 'Manure')@otu_table, 'matrix'), 
            as(taxa_filter(phylo_object, treatments, subset = 'Manure-Treated Soil')@otu_table, 'matrix'))
wilcox.test(as(taxa_filter(phylo_object, treatments, subset = 'Soil')@otu_table, 'matrix'), 
            as(taxa_filter(phylo_object, treatments, subset = 'Manure-Treated Soil')@otu_table, 'matrix'))
wilcox.test(as(taxa_filter(phylo_object, treatments, subset = 'Soil')@otu_table, 'matrix'), 
            as(taxa_filter(phylo_object, treatments, subset = 'Manure-Treated Effluent')@otu_table, 'matrix'))
wilcox.test(as(taxa_filter(phylo_object, treatments, subset = 'Effluent')@otu_table, 'matrix'), 
            as(taxa_filter(phylo_object, treatments, subset = 'Manure-Treated Effluent')@otu_table, 'matrix'))
```

```{r anova-for-replicates}
dat <- taxa_filter(phylo_object, treatment = treatments, subset = 'Manure')
adonis(t(dat@otu_table) ~ Sample, data = as(dat@sam_data, 'data.frame'),  permutations=1000, method = "bray")

dat <- taxa_filter(phylo_object, treatment = treatments, subset = 'Soil')
adonis(t(dat@otu_table) ~ Sample, data = as(dat@sam_data, 'data.frame'),  permutations=1000, method = "bray")

dat <- taxa_filter(phylo_object, treatment = treatments, subset = 'Effluent')
adonis(t(dat@otu_table) ~ Day * Sample, data = as(dat@sam_data, 'data.frame'),  permutations=1000, method = "bray")

dat <- taxa_filter(phylo_object, treatment = treatments, subset = 'Manure-Treated Soil')
adonis(t(dat@otu_table) ~ Day * Sample, data = as(dat@sam_data, 'data.frame'),  permutations=1000, method = "bray")

dat <- taxa_filter(phylo_object, treatment = treatments, subset = 'Manure-Treated Effluent')
adonis(t(dat@otu_table) ~ Day * Sample, data = as(dat@sam_data, 'data.frame'),  permutations=1000, method = "bray")
```

#Identifying Persisting Genes from Manure

```{r identify-persisters}
sum(unique_control_genes$Manure %in% taxa_names(taxa_filter(phylo_object, 'Treatment', 'Manure')))
sum(unique_control_genes$Manure %in% taxa_names(taxa_filter(phylo_object, treatments, 'Manure-Treated Effluent')))
manure_persisters_genes <- unique_control_genes$Manure[unique_control_genes$Manure %in% taxa_names(taxa_filter(phylo_object, 'Treatment', 'Manure', frequency = 0.5))]
persisters_in_manure <- taxa_filter(subset_taxa(phylo_object, taxa_names(phylo_object) %in% manure_persisters_genes), treatment = treatments, subset = 'Manure', drop_samples = TRUE)
sum(taxa_sums(persisters_in_manure)) / sum(taxa_sums(taxa_filter(phylo_object, treatments, subset = 'Manure')))
manure_persisters <- taxa_filter(subset_taxa(phylo_object, taxa_names(phylo_object) %in% manure_persisters_genes), treatment = 'Treatment', subset = 'Manure', drop_samples = TRUE)
```

```{r all-persister-merged, fig.width=16, fig.height=12}
manure_persisters <- conglomerate_samples(manure_persisters, treatments, merge_on = 'Day')
```

```{r manure-persister-composition}
data.frame(taxa_proportions(subset_taxa(manure_persisters, 'Matrix_Treatment' != 'Manure'), 'ARG_Class', treatments))
```

```{r manure-persister-composition-profile}
phylogeny_profile(manure_persisters, classification = 'ARG_Class', treatment = treatments, relative_abundance = FALSE) + 
  labs(color = 'Associated Anti-Microbial\nResistance Drug Class', fill = 'Associated Anti-Microbial\nResistance Drug Class', x = 'Sampling Day')
```

```{r all-persister-heatmaps-merged, fig.width=16, fig.height=12}
graph_data <- melt_phyloseq(manure_persisters)
colors <- phylosmith:::create_palette(length(unique(graph_data[['ARG_Class']])), 'default')
abundance_heatmap(manure_persisters, classification = 'Gene', treatment = treatments, transformation = 'log') + 
  theme(axis.text.y = element_text(colour = colors, face = 'bold')) + labs(color = 'Natural Log of\nGene Abundance', fill = 'Natural Log of\nGene Abundance', x = 'Sampling Day')
```

```{r all-persister-lines, fig.width=16, fig.height=8}
abundance_lines(manure_persisters, 'Gene', treatments) + scale_x_discrete(expand = expand_scale(mult = c(0,0.02))) + 
  theme(panel.spacing = unit(3, "cm")) + guides(colour = guide_legend(ncol = 1, override.aes = list(size = 4))) + 
  labs(color = 'Gene', x = 'Sampling Day')
```

#Identify Enriched Genes

```{r identify-stimulated-lists}
Soil <- taxa_filter(phylo_object, treatments, subset = 'Soil', frequency = 0)
Soil_genes <- taxa_names(Soil)
effluent <- taxa_filter(phylo_object, treatments, subset = 'Effluent', frequency = 0)
effluent_genes <- taxa_names(effluent)
manure_treated_soil <- taxa_filter(phylo_object, treatments, subset = 'Manure-Treated Soil')
manure_treated_soil_genes <- taxa_names(manure_treated_soil)

water_stimulated <- effluent_genes[!(effluent_genes %in% Soil_genes)]
water_stimulated <- water_stimulated[!(water_stimulated %in% manure_genes)]; length(water_stimulated)
water_enhanced <- effluent_genes[effluent_genes %in% Soil_genes]
water_enhanced <- water_enhanced[taxa_sums(subset_taxa(effluent, taxa_names(effluent) %in% water_enhanced))/nsamples(effluent) > taxa_sums(subset_taxa(Soil, taxa_names(Soil) %in% water_enhanced))/nsamples(Soil)]; length(water_enhanced)
water_enriched <- c(water_enhanced, water_stimulated); length(water_enriched)

manure_stimulated <- manure_treated_soil_genes[!(manure_treated_soil_genes %in% soil_genes)]
manure_stimulated <- manure_stimulated[!(manure_stimulated %in% manure_genes)]
length(manure_stimulated)
manure_enhanced <- manure_treated_soil_genes[manure_treated_soil_genes %in% soil_genes]
manure_enhanced <- manure_enhanced[taxa_sums(subset_taxa(manure_treated_soil, taxa_names(manure_treated_soil) %in% manure_enhanced))/nsamples(manure_treated_soil) > taxa_sums(subset_taxa(soil, taxa_names(soil) %in% manure_enhanced))/nsamples(soil)]
manure_enhanced <- manure_enhanced[!(manure_enhanced %in% water_enriched)]; length(manure_enhanced)
```

```{r gene-labeling, echo=FALSE}
manure_persisters_genes <- unique_control_genes$Manure[unique_control_genes$Manure %in% taxa_names(taxa_filter(phylo_object, 'Treatment', 'Manure', frequency = 0))]
gene_classification <- list(
  Manure_Persister = manure_persisters_genes,
  Soil = c(soil_genes[!(soil_genes %in% c(water_enriched, manure_enhanced))]),
  Water_Enriched = water_enriched,
  Manure_Enriched = manure_enhanced
)
Background = taxa_names(phylo_object)[!(taxa_names(phylo_object) %in% unlist(gene_classification))]
gene_classification$Background <- Background

gene_classification <- data.table(Classification = rep(names(gene_classification), sapply(gene_classification, length)),
  gene = unlist(gene_classification))
gene_classification <- gene_classification[match(rownames(phylo_object@tax_table), gene_classification$gene), 1]
phylo_object@tax_table <- tax_table(cbind(phylo_object@tax_table, as.matrix(gene_classification)))
```

```{r manure-persister-genes}
manure_persisters_genes
```

```{r gene-profile}
new <- as(phylo_object@tax_table, 'matrix')
new <- new[,c(6,4)]
newnew <- phylo_object
newnew@tax_table <- tax_table(new)
newnew <- conglomerate_samples(newnew, treatment = treatments)
new <- conglomerate_taxa(newnew, 'ARG_Class')

graph_data <- melt_phyloseq(taxa_filter(new, treatment = 'Matrix_Treatment', subset = 'Manure-Treated Effluent'))
graph_data$Classification <- factor(graph_data$Classification, levels = c('Water_Enriched', 'Manure_Enriched', 'Manure_Persister', 'Soil', 'Background'))
levels(graph_data$Classification) <- c('Water Enriched', 'Manure Enriched', 'Manure Persister', 'Soil', 'Background')
# graph_data[, Relative_Abundance := Abundance/sum(Abundance), by = Classification]
graph_data <- graph_data[!(graph_data$Classification %in% c('Soil', 'Background')), ]
color_count <- length(unique(graph_data[['ARG_Class']]))
graph_colors <- phylosmith:::create_palette(color_count)


ggplot(graph_data, aes_string(x = "Classification", y = "Abundance", 
                                   fill = 'ARG_Class')) + 
  guides(fill = guide_legend(ncol = ceiling(length(unique(graph_data[['ARG_Class']]))/25)), color = FALSE) + 
  scale_fill_manual(values = graph_colors, aesthetics = c("color", "fill")) + 
  facet_grid(reformulate('Classification'), scales = "free", space = "free") + 
  geom_bar(aes_string(color = 'ARG_Class', fill = 'ARG_Class'), stat = "identity", position = "stack", size = 0.8, color = 'black') + 
  ylab('Abundance') + 
  theme_classic() + theme(text = element_text(family = "times"),
                                axis.text.x = element_blank(), axis.text.y = element_text(size = 10), axis.title.x = element_blank(), axis.title.y = element_text(size = 10, 
                                                                                                                                                            face = "bold"), axis.ticks.x = element_blank(), 
                          legend.title = element_text(size = 10, face = "bold"), 
                          legend.text = element_text(size = 8), 
                          legend.spacing.x = unit(0.01, "npc"), 
                          panel.background = element_rect(color = "black", 
                                                          size = 1.5, fill = "white"), panel.spacing = unit(0.01, "npc"), strip.text.x = element_text(size = 10, face = "bold"), 
                          strip.background = element_rect(colour = "black", size = 1.4),
                          legend.key.size = unit(4, "mm")) + 
  scale_y_continuous(expand = expand_scale(mult = c(0.0037, 0.003), add = c(0, 0))) + 
  scale_x_discrete(expand = expand_scale(mult = 0, add = 0.51)) + 
labs(color = 'Associated Anti-Microbial\nResistance Drug Class', fill = 'Associated Anti-Microbial\nResistance Drug Class', x = 'Sampling Day')
```


```{r gene-proportions-manure-effluet}
taxa_proportions(phylo_object, classification = 'Classification', treatment = treatments)[Matrix_Treatment == 'Manure-Treated Effluent']
```


```{r gene-proportions-effluent}
taxa_proportions(phylo_object, classification = 'Classification', treatment = treatments)[Matrix_Treatment == 'Effluent']
```

#Co-Occurrence

```{r comput-significant-rho, cache=TRUE, eval=FALSE}
rho_cutoff <- permute_rho(phylo_object, treatments, replicate_samples = 'Day', permutations = 1000)
quantile_permuted_rhos(rho_cutoff)
```

```{r co-occurrence}
co_occurrence_table <- co_occurrence(phylo_object, treatments, p = 0.05)
co_occurrence_table$p <- p.adjust(co_occurrence_table$p, "BH")
co_occurrence_table <- co_occurrence_table[p < .05]
co_occurrence_table <- co_occurrence_table[rho <= -0.68 | rho >= 0.86]
```

```{r effluent_networks_gene}
layout <- network_layout_ps(phylo_object, treatments, 'Manure-Treated Effluent', co_occurrence_table, algorithm = 'fr')
a <- co_occurrence_network(phylo_object, 'Classification', treatments, 'Manure-Treated Effluent', co_occurrence_table = co_occurrence_table, layout = layout, cluster = TRUE) + 
  labs(color = 'Gene Behavior', fill = 'Gene Behavior')

b <- co_occurrence_network(phylo_object, 'ARG_Class', treatments, 'Manure-Treated Effluent', co_occurrence_table, layout = layout, cluster = TRUE) + 
  labs(color = 'Associated Anti-Microbial\nResistance Drug Class', fill = 'Associated Anti-Microbial\nResistance Drug Class')

grid.arrange(b, a, ncol = 2)
```

```{r effluent_networks_mechanism}
co_occurrence_network(phylo_object, 'ARG_Mechanism', treatments, 'Manure-Treated Effluent', co_occurrence_table, layout = layout, cluster = TRUE, nodes_of_interest = c('tet(33)', 'ermB'))
```